embedplyr: Tools for Working With Text Embeddings

### 2025-01-14 README.Rmd
minor revisions

```{r, eval=FALSE}
# function that takes character vector and outputs a data frame
sbert_embeddings <- function(texts) {
	text::textEmbed(
		texts,
		model = "sentence-transformers/all-MiniLM-L12-v2", # model name
		layers = -2,  # second to last layer (default)
		tokens_select = "[CLS]", # use only [CLS] token
		dim_name = FALSE,
		keep_token_embeddings = FALSE
	)$texts[[1]]
}

valence_sbert_df <- valence_df |> 
	embed_docs("text", sbert_embeddings, id_col = "id", .keep_all = TRUE)
```

>>>>

```{r, eval=FALSE}
# add embeddings to data frame
valence_sbert_df <- valence_df |> 
	embed_docs(
		"text", text::textEmbed, id_col = "id", .keep_all = TRUE,
		model = "sentence-transformers/all-MiniLM-L12-v2"
		)
```

### 2025-01-14 paper/paper.Rmd
minor revisions

The following example embeds three texts, which for the sake of the example I will consider to have been written by a participant diagnosed with depression, one diagnosed with anxiety, and one control.

@@

I use Distributed Dictionary Representation (DDR) as a showcase for **embedplyr**'s functionality. DDR enables the application of validated psychometric lexicons [e.g. @boyd2022] to rich, embedding-based semantic representation [@garten2018]. This is achieved by retrieving pretrained word embeddings for each word in the dictionary, and averaging them to create a single vector---the DDR. The dictionary construct can then be measured by comparing text embeddings to the DDR.

@@

In the following example, I construct DDRs for high and low anxiety. For the sake of the example, I use made-up dictionaries. Once embeddings are produced for each word in the dictionary, they are averaged using `average_embedding()`. By default this is a simple mean, but `average_embedding()` also supports the geometric median [@cardot2022], weighted geometric median, and weighted mean, including weighting by word frequency or smooth inverse frequency [@arora2017].

@@

To complete the DDR analysis, I compare the embeddings of each the corpus texts to that of the DDR. This could be done by computing cosine similarity between each text and `high_anx_DDR` (`"cosine"` is the default method for `get_sims()`). In this case however, I use an anchored vector to quantify the extent to which these texts reflect high anxiety _as opposed to low anxiety_. `method = "anchored"` gives the position of each embedding on the spectrum between two anchor points, where vectors aligned with `pos` are given a score of 1 and those aligned with `neg` are given a score of 0. This approach is also known as semantic projection [@grand2022].

@@

### Loading Other Embeddings

@@

```{r, eval=FALSE}
# add embeddings to data frame
valence_sbert_df <- valence_df |> 
	embed_docs("text", text::textEmbed, id_col = "id", .keep_all = TRUE)
```

>>>>

The following example embeds three texts, which for the sake of the example will can considered to have been written by one participant diagnosed with depression, one diagnosed with anxiety, and one control.

@@

Distributed Dictionary Representation (DDR) enables the application of validated psychometric lexicons [e.g. @boyd2022] to rich, embedding-based semantic representation [@garten2018]. This is achieved by retrieving pretrained word embeddings for each word in the dictionary, and averaging them to create a single vector---the DDR. The dictionary construct can then be measured by comparing text embeddings to the DDR.

@@

In the following example, DDRs are constructed for high and low anxiety using example dictionaries. Once embeddings are produced for each word in the dictionary, they are averaged using `average_embedding()`. By default this is a simple mean, but `average_embedding()` also supports the geometric median [@cardot2022], weighted geometric median, and weighted mean, including weighting by word frequency or smooth inverse frequency [@arora2017].

@@

To complete the DDR analysis initiated above, the embeddings of each the corpus texts are compared to that of the DDR. This could be done by computing cosine similarity between each text and `high_anx_DDR` (`"cosine"` is the default method for `get_sims()`). In this case however, an anchored vector is used to quantify the extent to which these texts reflect high anxiety _as opposed to low anxiety_. `method = "anchored"` gives the position of each embedding on the spectrum between two anchor points, where vectors aligned with `pos` are given a score of 1 and those aligned with `neg` are given a score of 0. This approach is also known as semantic projection [@grand2022].

@@

### Read Embeddings From File

@@

```{r, eval=FALSE}
# add embeddings to data frame
valence_sbert_df <- valence_df |> 
	embed_docs(
		"text", text::textEmbed, id_col = "id", .keep_all = TRUE,
		model = "sentence-transformers/all-MiniLM-L12-v2"
		)
```
